---
title: "Weather Patterns in 5 Major Cities"
author: "Alliyah Galvan & Jason Nguyen"
date: "March 13, 2020"
output: ioslides_presentation
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library("xml2")
library("glue")
library("skimr")
library("tidyverse")
library("naniar")
library("shiny")
library("shinydashboard")
```

```{r include=FALSE}
weather <- readr::read_csv("data/Weather.csv")
```

```{r include=FALSE}
new_factor<- c(1:5,26)
new_numeric<- c(6:25)
weather[,new_factor] <- lapply(weather[,new_factor], as.factor)
weather[,new_numeric] <- lapply(weather[,new_numeric], as.numeric)
lapply(weather, class) #changing classes
```

## Background | Data
-  <font size="4"> Our data was provided by: https://vincentarelbundock.github.io/Rdatasets/doc/mosaicData/Weather.html </font>
- <font size="4"> The data is an average of different weather patterns collected every day from 2016-2017 in 5 major cities: Auckland, New Zealand | Beijing, China | Chicago, USA | Mumbai, India | San Diego, USA</font>
- <font size="4"> Weather patterns include: Temperature, Dewpoint, Humidity, Wind, and Weathering Events</font>

## Shiny App

```{r include=FALSE}
weather_temp <- weather %>% 
  select(year, city, avg_temp) %>% 
  group_by(city, year) %>%
  summarize(Average_Temp=mean(avg_temp)) %>%
  rename(City="city", Year="year")
```

```{r include=FALSE}
weather_dewpt <-weather %>% 
  select(year, city, avg_dewpt) %>% 
  group_by(city, year) %>%
  summarize(Average_Dew_Point=mean(avg_dewpt)) %>% 
  select(-c("year"))
```

```{r include=FALSE}
weather_humidty <- weather %>% 
  select(year, city, avg_humidity) %>% 
  group_by(city, year) %>%
  summarize(Average_Humidty=mean(avg_humidity)) %>%
  select(-c("year"))
```

```{r include=FALSE}
weather_wind <- weather %>% 
  select(year, city, avg_wind) %>% 
  group_by(city, year) %>%
  summarize(Average_Wind=mean(avg_wind)) %>%
  select(-c("year"))
```

```{r results= 'hide', character}
weather_mean <- cbind(weather_temp, weather_dewpt, weather_humidty, weather_wind) #combining yearly average into one table
weather_cleaner <- weather_mean %>% 
  select(-c(4,6,8)) %>% 
  pivot_longer(Average_Temp:Average_Wind,
               names_to = "Measure",
               values_to = "Value")
weather_cleaner
```

- The Shiny App allows for a visual representation of the yearly comparisons between each city for temperature, dewpoint, humidity, and wind.
- We had to make multiple tables of each category and combine them. Then, tidy the data using pivot longer.
- The Shiny App is present in its own file in our Repository.

## **Average Monthly Temperatures for All Cities for 2016**
```{r echo=FALSE, out.width = '60%', fig.align = "center"}
weather %>% 
  mutate_at(vars(month), as.factor) %>% 
 filter(year== 2016) %>%
  group_by(month, city) %>% 
  summarize(mean_avgtemp=mean(avg_temp))  %>% 
  ggplot(aes(x=month, y=mean_avgtemp, color=city, group=city))+
  geom_line() +
  geom_point(alpha=0.8)+
  labs(title="2016 Monthly Average Temperature",
       x= "Month",
       y="Average Temperature (Fº)") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

- <font size="2">To break things down further, here is a look at the average temperature for each city in 2016.</font>
- <font size="2">It is evident that Auckland lies in the Southern Hemisphere while the other cities lie in the Northern Hemisphere.</font>
- <font size="2">Auckland increases temperature while the other cities' average temperatures decrease and vice versa.</font>

## **Average Monthly Temperatures for All Cities for 2017**
```{r echo=FALSE, out.width = '60%', fig.align = "center"}
weather %>% 
  mutate_at(vars(month), as.factor) %>% 
 filter(year== 2017) %>%
  group_by(month, city) %>% 
  summarize(mean_avgtemp=mean(avg_temp))  %>% 
  ggplot(aes(x=month, y=mean_avgtemp, color=city, group=city))+
  geom_line() +
  geom_point(alpha=0.8)+
  labs(title="2017 Monthly Average Temperature",
       x= "Month",
       y="Average Temperature (Fº)") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

- <font size="2">Similar temperature patterns are shown in 2017 with slight fluctuations.</font>
- <font size="2">We also notice that  Mumbai has overall the highest temperature because it is closest to the equator.</font>

## **Average Monthly Dewpoints for All Cities for 2016**
```{r echo=FALSE, out.width = '60%', fig.align = "center"}
weather %>% 
  mutate_at(vars(month), as.factor) %>% 
 filter(year== 2016) %>%
  group_by(month, city) %>% 
  summarize(mean_avgdewpt=mean(avg_dewpt))  %>% 
  ggplot(aes(x=month, y=mean_avgdewpt, color=city, group=city))+
  geom_line() +
  geom_point(alpha=0.8)+
  labs(title="2016 Monthly Average Dewpoint %",
       x= "Month",
       y="Average Dewpoint %") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

- <font size="2">Average dewpoint percentages show similar results to average temperatures as dewpoint is defined as the measure of how much moisture is in the air.</font>

## **Average Monthly Dewpoints for All Cities for 2017**
```{r echo=FALSE, out.width = '60%', fig.align = "center"}
weather %>% 
  mutate_at(vars(month), as.factor) %>% 
 filter(year== 2017) %>%
  group_by(month, city) %>% 
  summarize(mean_avgdewpt=mean(avg_dewpt))  %>% 
  ggplot(aes(x=month, y=mean_avgdewpt, color=city, group=city))+
  geom_line() +
  geom_point(alpha=0.8)+
  labs(title="2017 Monthly Average Dewpoint %",
       x= "Month",
       y="Average Dewpoint %") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

- Again here is the overall year of 2017 showing the dewpoint monthly average percentages for each of the 5 cities.

## **Average Monthly Humidity for All Cities for 2016**
```{r echo=FALSE, out.width = '60%', fig.align = "center"}
weather %>% 
  mutate_at(vars(month), as.factor) %>% 
 filter(year== 2016) %>%
  group_by(month, city) %>% 
  summarize(mean_avghumidity=mean(avg_humidity))  %>% 
  ggplot(aes(x=month, y=mean_avghumidity, color=city, group=city))+
  geom_line() +
  geom_point(alpha=0.8)+
  labs(title="2016 Monthly Average Humidity %",
       x= "Month",
       y="Average Humidity %") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

- <font size="2"> Humidity varies from the two previous graphs. As seen, Auckland has a consistently high humidity percentage while the other cities have greater fluctuation.</font> 
- <font size="2">This is also in fact due to the fact that Auckland lies closest to the equator and does not vary in weather patterns as much.</font>

## **Average Monthly Humidity for All Cities for 2017**
```{r echo=FALSE, out.width = '60%', fig.align = "center"}
weather %>% 
  mutate_at(vars(month), as.factor) %>% 
 filter(year== 2017) %>%
  group_by(month, city) %>% 
  summarize(mean_avghumidity=mean(avg_humidity))  %>% 
  ggplot(aes(x=month, y=mean_avghumidity, color=city, group=city))+
  geom_line() +
  geom_point(alpha=0.8)+
  labs(title="2017 Monthly Average Humidity %",
       x= "Month",
       y="Average Humidity %") +
  theme(plot.title = element_text(size = 18, face = "bold"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12))
```

- In 2017, the overall humidity greatly resembles that of 2016 for these major cities.

## **Chaning NA's for Events Column**
```{r echo=TRUE}
weather$events<- as.character(weather$events)
weather$events[is.na(weather$events)] <- "Clear Skies"
skies <- weather %>% 
  select(city, events)
```

- We had NA's in the Weathering Events Column. But since we wanted to find out the types of weather events each city had, we changed the NA into "Clear Skies".

## **Auckland Weather Events**

```{r echo=FALSE, out.width = '70%', fig.align = "center"}
skies %>%
  filter(city== "Auckland") %>% 
  ggplot(aes(x=events)) +
  geom_bar(fill = "brown1", color="black", alpha =.5) +
  labs(title = "Weather Events in Auckland",
       x= "Weather Events",
       y= "Count") +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),)+
  coord_flip()
```

- Auckland's weathering events are a bit of a surprise; rain takes precedence over clear skies.

## **Beijing Weather Events**
```{r echo=FALSE, out.width = '70%', fig.align = "center"}
skies %>%
  filter(city== "Beijing") %>% 
  ggplot(aes(x=events)) +
  geom_bar(fill = "darkgoldenrod2", color="black", alpha =.5) +
  labs(title = "Weather Events in Beijing",
       x= "Weather Events",
       y= "Count") +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),)+
  coord_flip()
```

- Beijing's full of days with clear skies within these two years. Rain trails along as the next common weathering event.

## **Chicago Weather Events**
```{r echo=FALSE, out.width = '70%', fig.align = "center"}
skies %>%
  filter(city== "Chicago") %>% 
  ggplot(aes(x=events)) +
  geom_bar(fill = "palegreen1", color="black", alpha =.5) +
  labs(title = "Weather Events in Chicago",
       x= "Weather Events",
       y= "Count") +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),)+
  coord_flip()
```

- As expected, Chicago shows variation throughout the two years with a variety of different patterns. However, Chicago still maintains its days with clear skies as higher than any other weathering event.

## **Mumbai Weather Events**
```{r echo=FALSE, out.width = '70%', fig.align = "center"}
skies %>%
  filter(city== "Mumbai") %>% 
  ggplot(aes(x=events)) +
  geom_bar(fill = "deepskyblue", color="black", alpha =.5) +
  labs(title = "Weather Events in Mumbai",
       x= "Weather Events",
       y= "Count") +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),)+
  coord_flip()
```

- For a majority of the year, Mumbai shows clear skies. Rainy days are not as common as clear skies but appear more frequently than any other weather event or combinations of events.

## **San Diego Weather Events**
```{r echo=FALSE, out.width = '70%', fig.align = "center"}
skies %>%
  filter(city== "San Diego") %>% 
  ggplot(aes(x=events)) +
  geom_bar(fill = "magenta", color="black", alpha =.5) +
  labs(title = "Weather Events in San Diego",
       x= "Weather Events",
       y= "Count") +
  theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),)+
  coord_flip()
```

- San Diego as predicted has a majority of clear skies throughout the two years of data. 
- They don't call it "Sunny San Diego" for nothing!